---
title: "STAT 209 Project 1"
author: "Group 1"
output: beamer_presentation

---

We are using the `Marraige` dataset provided in R package `mosaicData`. It provides marriage records from the Mobile County, AL probate court between 11/9/96 and 2/6/99.  Variables recorded include:

* **age - age of person**
* **college - number of years in college**
* **person - a factor with levels Bride and Groom**
* **prevconc - way the last marriage ended**
* **prevcount - number of previous marriages**
* **race - race of person, being American Indian, Black, Hispanic, and White**


```{r message = FALSE, echo = FALSE}
library(tidyverse)
data(Marriage, package = "mosaicData")
library(ggthemes)
```
---
```{r message = FALSE, echo = FALSE}
marriage_plot <- ggplot(Marriage, aes(x = prevconc, y = age)) + geom_boxplot(aes(color = person), fill = "aliceblue") + theme_economist()
marriage_plot <- marriage_plot + ggtitle("Comparison of How Prvious Marriage Ended Based on Age") +
  xlab("Reason for Ending Previous Marriage") +
  ylab("Age")
marriage_plot
```
---
```{r message = FALSE, echo = FALSE}
#Creates modified data set that filters out the entries for American Indian and Hispanic and creates a collumn for total years in school (high school + college)
Marriage.modified <- Marriage %>% filter(race %in% c("White", "Black"))
Marriage.modified <- Marriage.modified %>% mutate(schooling = hs + college)
```

---

```{r message = FALSE, echo = FALSE}
#Julian
ggplot(Marriage.modified, aes(x= as.factor(prevcount), y = delay)) +
  facet_wrap(~race)+
  #geom_point()+
  geom_boxplot(aes(color = race), fill = "aliceblue") +
  ggtitle("Time between Marriage Application and Ceremony Date") +
  ylab("Ceremony Delay") +
  xlab("Previous Times Married") +
  theme_economist() +
  theme(axis.text = element_text(size = 16), 
        axis.title = element_text(size = 20))
```
---
```{r message = FALSE, echo = FALSE}
race_data <- filter(Marriage, race == c("White","Black"))

tally_count <- race_data %>% group_by(college) %>% tally()
my_plot <- race_data %>%
  group_by(college,race) %>%
  summarize(r_count = n()) %>%
  ggplot(aes(x = college, y = r_count, group=race, color = race)) + geom_bar(stat = "identity", fill="aliceblue")
edu_bar <-my_plot + facet_wrap(~race) + theme_economist() +ggtitle("Education When Getting Married") + ylab("Number of People") + xlab("Years of College")
edu_bar + +theme(axis.text=element_text(size=12), axis.title=element_text(size=14,face="bold"))
```
