---
title: "STAT 209 Project 1"
author: "Jared Steinberg, Dizzy Farbanish, Julian Grabarek, Luis Solorzano"
output: beamer_presentation
---

We are using the `Marraige` dataset provided in R package `mosaicData`. It provides marriage records from the Mobile County, AL probate court between 11/9/96 and 2/6/99.  Variables recorded include:

* **age - age of person**
* **college - number of years in college**
* **person - a factor with levels Bride and Groom**
* **prevconc - way the last marriage ended**
* **prevcount - number of previous marriages**
* **race - race of person, being American Indian, Black, Hispanic, and White**


```{r message = FALSE, echo = FALSE}
library(tidyverse)
data(Marriage, package = "mosaicData")
library(ggthemes)
```




#Graph 1
```{r message = FALSE, echo = FALSE}
marriage_plot <- ggplot(Marriage, aes(x = prevconc, y = age)) + geom_boxplot(aes(color = person), fill = "aliceblue") + theme_economist()
marriage_plot <- marriage_plot + ggtitle("Comparison of How Previous Marriage Ended Based on Age") +
  xlab("Reason for Ending Previous Marriage") +
  ylab("Age") +
  theme(axis.text = element_text(size = 16), 
        axis.title = element_text(size = 20))+
  theme(plot.title = element_text(size = 22, face = "bold")) +
  theme(plot.title = element_text(hjust = 0.5))
marriage_plot
```

----

```{r message = FALSE, echo = FALSE}
#Creates modified data set that filters out the entries for American Indian and Hispanic and creates a collumn for total years in school (high school + college)
Marriage.modified <- Marriage %>% filter(race %in% c("White", "Black"))
Marriage.modified <- Marriage.modified %>% mutate(schooling = hs + college)
```


```{r message = FALSE, echo = FALSE}
#Julian
delay_plot <- ggplot(Marriage.modified, aes(x= as.factor(prevcount), y = delay)) +
  facet_wrap(~race)+
  #geom_point()+
  geom_boxplot(aes(color = race), fill = "aliceblue") +
  ggtitle("Time between Marriage Application and Ceremony Date") +
  ylab("Ceremony Delay") +
  xlab("Previous Times Married") +
  theme_economist() +
  theme(axis.text = element_text(size = 16), 
        axis.title = element_text(size = 20)) +
  theme(plot.title = element_text(size = 22, face = "bold")) +
  theme(plot.title = element_text(hjust = 0.5))
delay_plot
```

----

```{r message = FALSE, echo = FALSE, warning= FALSE}
race_data <- filter(Marriage, race == c("White","Black"))

tally_count <- race_data %>% group_by(college) %>% tally()
my_plot <- race_data %>%
  group_by(college,race) %>%
  summarize(r_count = n()) %>%
  ggplot(aes(x = college, y = r_count, group=race, color = race)) + geom_bar(stat = "identity", fill="aliceblue")
edu_bar <-my_plot + facet_wrap(~race) + theme_economist() +ggtitle("Education When Getting Married") + ylab("Number of People") + xlab("Years of College")+
  theme(axis.text = element_text(size = 16), 
        axis.title = element_text(size = 20)) +
  theme(plot.title = element_text(size = 22, face = "bold")) +
  theme(plot.title = element_text(hjust = 0.5))
edu_bar
```

----

```{r message = FALSE, echo = FALSE}
dizzy_plot <- ggplot(Marriage, aes(y = prevcount, x = age, color = race)) + geom_smooth(fill = "aliceblue") + theme_economist() + ylab("Previous Divorces") + xlab("Age") +
  ggtitle("Comparison of Age vs. Number of Divorces") +
  theme(axis.text = element_text(size = 16), 
        axis.title = element_text(size = 20)) +
  theme(plot.title = element_text(size = 22, face = "bold"))+
  theme(plot.title = element_text(hjust = 0.5))
#+geom_spline
dizzy_plot
```


<!-- #Data shows that older generations have actually had less divorces than younger generations. This suggests that societal norms surrounding divorce have changed from the older to the younger generations. It is expected that the divorces averages should trend up with age but it is unexpected that divorce would be lower for older people from previous generations. Additionally it is interesting to note that divorces among african americans are lower on average than white people. The difference is statistically significant at 40. -->

<!-- Jared's paragraph:
My graph illustrates how reasons for ending previous marriages change by age.  I also subdivided my boxplots by gender, with grooms being blue boxplots and brides being pink boxplots.  The data shows that, for death being the reason the previous marriage ended, women’s age vary more than men’s, however have a lower median age.  For divorcees, both intervals for men and women are approximately the same size and median age, probably because people in a relationship are usually the same age.  The N/A category shows that people who have never had a previous marriage are typically young, regardless of sex.
-->

