---
title: "STAT 209 Project 1"
author: "Group 1"
output:
  html_document: default

---

We are using the `Marraige` dataset provided in R package `mosaicData`. It provides marriage records from the Mobile County, AL probate court between 11/9/96 and 2/6/99.  Variables recorded include:

* **age - age of person**
* **college - number of years in college**
* **person - a factor with levels Bride and Groom**
* **prevconc - way the last marriage ended**
* **race - race of person, being American Indian, Black, Hispanic, and White**


```{r message = FALSE, echo = FALSE}
library(tidyverse)
data(Marriage, package = "mosaicData")
library(ggthemes)
```


```{r}
marriage_plot <- ggplot(Marriage, aes(x = prevconc, y = age)) + geom_boxplot(aes(color = person), fill = "aliceblue") + theme_economist()
marriage_plot <- marriage_plot + ggtitle("Comparison of How Prvious Marriage Ended Based on Age")
marriage_plot
```

```{r}
#Creates modified data set that filters out the entries for American Indian and Hispanic and creates a collumn for total years in school (high school + college)
Marriage.modified <- Marriage %>% filter(race %in% c("White", "Black"))
Marriage.modified <- Marriage.modified %>% mutate(schooling = hs + college)
```

```{r}
ggplot(Marriage.modified, aes(x= as.factor(prevcount), y = delay)) +
  facet_wrap(~race)+
  #geom_point()+
  geom_boxplot(aes(color = race), fill = "aliceblue") +
  ggtitle("Time between Marriage Application and Ceremony Date") +
  ylab("Ceremony Delay") +
  xlab("Previous Times Married") +
  theme_economist()
```
